---
knitr: 
    opts_chunk: 
      code-fold: show
      results: hold
  
---
# Raster: Continuous {#sec-06}


```{r}
#| label: setup
#| results: hold
#| include: false

base::source(file = "R/helper.R")
ggplot2::theme_set(ggplot2::theme_bw())

```

## Table of content for chapter 04 {.unnumbered}

::::: {#obj-chap03}
:::: {.my-objectives}
::: {.my-objectives-header}
Chapter section list
:::

::: {.my-objectives-container}



:::
::::
:::::

## Importing Raster Data

### General data information

A raster object can be created by calling the `terra::rast()` function and specifying an external image file as an argument. … The resulting object belongs to the `SpatRaster` class. Invoking the `base::print()` function for a raster object provides information about the dimensions of the grid, cell size, geographic location, and other details.

::: {.my-code-collection}
:::: {.my-code-collection-header}
::::: {.my-code-collection-icon}
:::::
:::::: {#exm-06-get-general-data-information}
: Get general sparial raster data information
::::::

::::

::::{.my-code-collection-container}

::: {.panel-tabset}

###### rast()

:::::{.my-r-code}
:::{.my-r-code-header}
:::::: {#cnj-06-lst-example}
: Loading Land Survey Temperature (LST) Data
::::::
:::
::::{.my-r-code-container}
```{r}
#| label: lst-example

lst <- terra::rast("data/Chapter6/MOD11A2_2017-07-12.LST_Day_1km.tif")
base::cat("*********** class of data *************\n")
class(lst)
base::cat("\n**************** metadata *****************\n")
lst

```

***

**From the help file:**

`terra::rast()` is the method to create `SpatRaster` objects. These objects can be created from scratch, from a filename, or from another object.

A `SpatRaster` represents a spatially referenced surface divided into three dimensional cells (rows, columns, and layers).

When a `SpatRaster` is created from one or more files, it does not load the cell (pixel) values into memory (RAM). It only reads the parameters that describe the geometry of the `SpatRaster`, such as the number of rows and columns and the coordinate reference system. The actual values will be read when needed.


::::
:::::



###### summary()

:::::{.my-r-code}
:::{.my-r-code-header}
:::::: {#cnj-06-summary-of-raster-data}
: Summary of Raster Data
::::::
:::
::::{.my-r-code-container}
```{r}
#| label: summary-of-raster-data

terra::summary(lst, size = 1e6)
```

***

The `terra::summary()` function provides information about the statistical distribution of raster values, with the `size` argument specifying the number of randomly sampled pixels to include in the summary.

::::
:::::

###### helpers

:::::{.my-r-code}
:::{.my-r-code-header}
:::::: {#cnj-06-helpers-function}
: Helpers function to extract specific characteristics of a SpatRaster objects
::::::
:::
::::{.my-r-code-container}
```{r}
#| label: helpers-function

base::cat("### Number of grid cells ###\n")
terra::ncell(lst)
base::cat("\n### Number of grid rows ###\n")
terra::nrow(lst)
base::cat("\n### Number of grid columns ###\n")
terra::ncol(lst)
base::cat("\n### Number of raster layers ###\n")
terra::nlyr(lst)
base::cat("\n###### Grid cell size ######\n")
terra::res(lst)
```



::::
:::::

###### ext()

:::::{.my-r-code}
:::{.my-r-code-header}
:::::: {#cnj-06-raster-extent}
: Return geographic coordinates of the raster extent
::::::
:::
::::{.my-r-code-container}
```{r}
#| label: raster-extend

lst_ext <- terra::ext(lst)

base::cat("###### Data class ######\n")
base::class(lst_ext)

base::cat("\n############ Raster coordinates ############\n")
lst_ext[1:4]

```

***

The `terra::ext()` function returns a `SpatExtent` object that contains the geographic coordinates of the raster extent. `SpatExtent` objects can be used to specify the extent of new raster objects or to crop rasters to a new extent.

::::
:::::

###### crs()

:::::{.my-r-code}
:::{.my-r-code-header}
:::::: {#cnj-06-get-crs}
: Get the coordinate reference system (CRS)
::::::
:::
::::{.my-r-code-container}
```{r}
#| label: get-crs

base::cat("###### Return details about the coordinate reference system ######\n")
terra::crs(lst, parse = TRUE)

base::cat("\n########## Return an abbreviated summary ##########\n")
terra::crs(lst, describe = TRUE)
```

The `terra::crs()` function returns a character string containing details about the coordinate reference system (`r glossary("CRS")`) of the raster dataset. By default, the output is in well-known text (`r glossary("WKT")`) format. The `parse = TRUE` argument parses the output into a vector that is easier to read when printed to the R console.

::::
:::::

###### names()

:::::{.my-r-code}
:::{.my-r-code-header}
:::::: {#cnj-06-get-set-layer-names}
: Get and set layer names
::::::
:::
::::{.my-r-code-container}
```{r}
#| label: get-set-layer-names

base::cat("###### get original layer name ######\n")
terra::names(lst)

base::cat("\n###### get changed layer name ######\n")
names(lst) <- c("temperature")
terra::names(lst)

```

::::
:::::


:::

::::
:::::

### Modifying raster data objects

`SpatRaster` objects can be modified with various functions and used as input to mathematical expressions. 

In the LST dataset, missing data are coded as zeroes, and the raw digital numbers must be modified by a scaling factor of 0.02 to convert them into degrees Kelvin. The code below uses the `terra::ifel()` function, which is analogous to the` base::ifelse()` function, to replace zero values in the LST dataset with `NA` values. 

The next expression applies the scaling factor to obtain degrees Kelvin and subtracts a constant to further convert the temperatures from Kelvin to Celsius. In both cases, the calculations are automatically carried out for every grid cell in the raster dataset.


:::::{.my-r-code}
:::{.my-r-code-header}
:::::: {#cnj-06-modifying-raster-data}
: Modiyfing raster data objects
::::::
:::
::::{.my-r-code-container}
```{r}
#| label: modifying-raster-data

lst <- terra::ifel(lst == 0, NA, lst)
lst_c <- lst * 0.02 - 273.15
terra::summary(lst_c, size = 1e6)
```

::::
:::::

### Statistical summaries

The `terra::global()` function can be used to generate statistical summaries of the pixels in a raster dataset. 

:::::{.my-r-code}
:::{.my-r-code-header}
:::::: {#cnj-06-statistical-summaries}
: Generate statistical saummaries with `terra::global()`
::::::
:::
::::{.my-r-code-container}
```{r}
#| label: statistical-summaries

terra::global(lst_c, fun = "mean", na.rm=T)
##                 mean
## temperature 30.28687
terra::global(lst_c, fun = "min", na.rm=T)
##              min
## temperature 1.67
terra::global(lst_c, fun = "max", na.rm=T)
##               max
## temperature 54.77
terra::global(lst_c, fun = "sd", na.rm=T)
##                   sd
## temperature 3.652403
```

::::
:::::

### Exporting raster objects

Raster objects can be exported using the `terra::writeRaster()` function. The format of the exported image is specified using the format argument. Common output formats are:

- GTiff (GeoTiff), 
- ascii (ESRI ASCII text ), 
- CDF, (NetCDF), see also `terra::writeCDF()` and 
- HFA (ERDAS Imagine). 

The `overwrite = TRUE` argument will replace existing files with the same name.

::: {.my-code-collection}
:::: {.my-code-collection-header}
::::: {.my-code-collection-icon}
:::::
:::::: {#exm-06-export-raster-objects}
: Exporting Raster Objects
::::::

::::

::::{.my-code-collection-container}

::: {.panel-tabset}

###### writeRaster()

:::::{.my-r-code}
:::{.my-r-code-header}
:::::: {#cnj-06-writeRaster-example}
: Exporting raster objects with `terra::writeRaster()`
::::::
:::
::::{.my-r-code-container}
<center>**Run this code chunk manually if the file still needs to be downloaded.**</center>

```{r}
#| label: writeRaster-example
#| eval: false

terra::writeRaster(lst_c, 
            filename = "data/Chapter6/MOD11A2_2017-07-12.LST_Day_1km_DegC.tif", 
            filetype = "GTiff", overwrite = TRUE )
```


::::
:::::


###### drivers

:::::{.my-r-code}
:::{.my-r-code-header}
:::::: {#cnj-06-get-gdal-drivers}
: Get available drivers for your system and gdal version
::::::
:::
::::{.my-r-code-container}
```{r}
#| label: get-gdal-drivers

terra::gdal(drivers = TRUE)
```



::::
:::::

:::

::::
:::::

***
