# Working with Austrian data {#sec-annex-d}


```{r}
#| label: setup
#| results: hold
#| include: false

base::source(file = "R/helper.R")
ggplot2::theme_set(ggplot2::theme_bw())
options(show.signif.stars = FALSE)

```


::::: {#obj-annex-d}
:::: {.my-objectives}
::: {.my-objectives-header}
Objectives
:::

::: {.my-objectives-container}

This chapter tries to find and to apply Austrian data for map plotting.

:::
::::
:::::

::: {.callout-warning #wrn-annex-d-austrian-data}
I can't remember from where I got the following Austrian data. Month later I tried to find the original source and learned about the following websites:

- [Gliederung Österreichs in Gemeinden](https://data.statistik.gv.at/web/meta.jsp?dataset=OGDEXT_GEM_1)
- [Update 2017 der Gemeindegrenzen](https://wahlen.strategieanalysen.at/geojson/)
- [Verwaltungsgrenzen Österreich](https://www.opendataportal.at/katalog/dataset/geojson-daten-osterreich) (GeoJSON, TopoJSON)
- [Download Austria Shapefiles](https://www.igismap.com/download-austria-shapefile-administrative-boundary-polygon-rail-road-line-ferry-airpoint-point/)
- [Austrian OGD basemap](https://basemap.at/en/)
- [Free Austria GIS Map files](https://simplemaps.com/gis/country/at#admin1) from [simplemaps](https://simplemaps.com/) 

***

The [Austrian Map](https://maps.bev.gv.at/#/center/13.35,47.7702/zoom/8) is an interesting website with all Austrian geographical and historical details. 
:::

:::::{.my-r-code}
:::{.my-r-code-header}
:::::: {#cnj-annex-d-load-regions}
: Load data for Austrian regions
::::::
:::
::::{.my-r-code-container}
::: {#lst-annex-d-load-regions}
```{r}
#| label: annex-d-load-regions

at_regions <- sf::st_read("_archive/austria/austria-with-regions_.geojson")
```

Load data for Austrian regions
:::

::::
:::::

:::::{.my-r-code}
:::{.my-r-code-header}
:::::: {#cnj-annex-d-plot-regions}
: Plot Austrian regions 
::::::
:::
::::{.my-r-code-container}
::: {#lst-annex-d-plot-regions}
```{r}
#| label: annex-d-plot-regions

ggplot2::ggplot(data = at_regions) +  
  ggplot2::geom_sf(fill = NA)
```

Austrian regions 
:::

::::
:::::

:::::{.my-r-code}
:::{.my-r-code-header}
:::::: {#cnj-annex-d-austrian-regions-with-admin-borders}
: Load Austrian regions with administrative borders
::::::
:::
::::{.my-r-code-container}
::: {#lst-annex-d-austrian-regions-with-admin-borders}
```{r}
#| label: annex-d-austrian-regions-with-admin-borders
#| cache: TRUE

at_folder = "_archive/austria"
pol_bez_folder = "/OGDEXT_POLBEZ_1_STATISTIK_AUSTRIA_20240101/"
pol_bez_file = "STATISTIK_AUSTRIA_POLBEZ_20240101.shp"

at_pol_bezirk <- sf::st_read(
  paste0(at_folder, pol_bez_folder, pol_bez_file)
)

ggplot2::ggplot(data = at_pol_bezirk) +  
  ggplot2::geom_sf(fill = NA)
```

Austrian regions with administrative borders
:::

::::
:::::


